\documentclass[aspectratio=1610]{beamer}
\usepackage{forest}
\usepackage{cgloss4e}
\usepackage{tikz}
\usepackage{gb4e}
\usetheme{Boadilla}
\usetikzlibrary{calc}
\usetikzlibrary{matrix}
\usetikzlibrary{positioning}

\title{Scope without Syntax}
\subtitle{Towards a Game Theoretic Approach}
\author{Luke Smith}
\institute{Committee: Robert, MPP, TgB}
\usetikzlibrary{shapes.geometric, arrows}

\tikzset{Above/.style={midway,above,font=\scriptsize,text width=1.5cm,align=center,},Below/.style={midway,below,font=\scriptsize,text width=1.5cm,align=center}}
\tikzstyle{box} = [rectangle, centered, draw=black,minimum width=3cm, minimum height=1cm]
\tikzstyle{arrow} = [thick,->,>=stealth]

\tikzset{centered/.style=}


\begin{document}

\resetcounteronoverlays{exx}

\begin{frame}
\titlepage
\end{frame}

\section{Background}

\begin{frame}
\frametitle{Quantifiers}\pause
\begin{itemize}
\item Languages have what are called \emph{quantifiers}, which are words which delineate particular quantities of nouns that they modify.\pause
\begin{itemize}
\item \textbf{Universal quantifiers} -- all, each, every ($\forall$)\pause
\item \textbf{Existential quantifiers} -- a, one, some ($\exists$)\pause
\item \textbf{Negation} -- not, no ($\neg$)\pause
\item Many others -- numerals, much, many, few, etc.\pause
\end{itemize}

\item For the purposes of sentence interpretation, quantifiers are quite a puzzle. Especially when there are multiple quantifiers in a sentence, a sentence may become ambiguous.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Scope Ambiguity}\pause
\begin{exe}
\ex Everyone loves someone.
\end{exe}\pause
\begin{itemize}
\item This sentence has two quantifiers, a universal ($\forall$) `every' and an existential ($\exists$) `some.'\pause
\item This sentence has two different interpretations:\pause
\begin{itemize}
\item For each person, there exists some other person  they love.\pause
\item There exists one particular person who everyone loves.\pause
\end{itemize}
\item In the first possible reading, we say that the $\forall$ takes `wide scope' over the $\exists$, which is said to have `narrow scope.'\pause

\item In the second, we say that the $\exists$ takes wide scope over the $\forall$.
\end{itemize}
\end{frame}

\section{Scope in the Field}

\begin{frame}
\frametitle{Traditional View}\pause
\begin{itemize}
\item Scope was traditionally dealt with in terms of `movement' and `logical form.' An ambiguous sentence had to go through some kind of post-syntactic change to yield an unambiguous representation in the mind.\pause
\item Different languages were discovered to have different availabilities of scope ambiguity. This was dealt with with formal and syntactic changes.\pause
\item Not so important to go into because basically nothing worked across wide data sets.\pause
\item Scope ambiguity is difficult to account for because it is:\pause
\begin{itemize}
\item Highly context sensitive (Chomsky's aphasia)\pause
\item Sensitive to linear order
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Game Theoretic Scope}\pause
\begin{itemize}
\item \textbf{My statement:} Scope ambiguity is totally paralinguistic. Scope ambiguities fall out from listeners' evaluation of the intentions of the speaker.\pause
\item This can partially be modeled in Game Theory, seeing that speakers are mutually evaluating the others' behavior and choosing how to word or interpret sentences based on that.\pause
\item This can allow us to formally analyze an apparent ``functional'' alternation.
\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{The Game}

\begin{figure}
\begin{center}
\begin{tikzpicture}[node distance=2cm]

\node (nature) [box, align=center] {{\large \textbf{ Player 0: ``Nature''}}\\Determines desired quantifier scope interpretation $s$};
\node (speaker) [box, align=center, below of=nature] {{\large \textbf{Player 1: ``Speaker''}}\\Determines a best wording $w$  to communicate $s$ };
\node (hearer) [box, align=center, below of=speaker] {{\large \textbf{Player 2: ``Hearer''}}\\Must guess $s$  based $w$};

\draw [arrow] (nature) -- (speaker) ;
\draw [arrow] (speaker) -- (hearer) ;
\end{tikzpicture}
\end{center}
\end{figure}
\end{frame}




\begin{frame}
\frametitle{Precedents in Linguistics}\pause
\begin{itemize}
\item Game Theory has been similarly employed in linguistics, particularly semantics to deal with implicatures.\pause
\begin{exe}
\ex Billy ate most of the chocolates.
\end{exe}\pause
\item Sentences like this in actual language are inferred to mean that Billy ate most \emph{but not all chocolates}, although the sentence is logically still true if he did.\pause
\item However speakers \emph{assume} Billy didn't eat \emph{all} the chocolates because if that were true, a speaker probably would've said so.
\end{itemize}
\end{frame}


\section{Potential Solution}

\begin{frame}
\frametitle{Assumptions and Constraints}
\begin{itemize}\pause
\item It is generally preferable if quantifiers occur in the order they are supposed to be interpreted in (surface scope).\pause
\item Moving around nouns via `transformations' (passivization, clefting, etc.) is costly/marked/undesirable.\pause
\item Scrambling (to be discussed later), as opposed to transformations are not similarly costly.
\end{itemize}
\end{frame}


\begin{frame}
\frametitle{English Data}\pause
\begin{itemize}
\item Typical English sentences show scope ambiguity if there is more than one quantifier:\pause
\begin{exe}
\ex Two men dug each hole.
\end{exe}\pause
\item There can be two particular men who dig all the holes ($\exists>\forall$)  or, each hole can be dug by a different pair of men ($\forall>\exists$) .\pause
\item Ambiguity will usually disappear or become highly dispreferred if the sentence undergoes a `transformation:'\pause
\begin{exe}
\ex Each hole was dug by two men.
\end{exe}\pause
\item Here, the strongly preferred reading is the one where there is a pair of men for each hole ($\forall>\exists$), while the case where there is two specific men for each hole is harder to get out of the blue.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{English Data}\pause
\begin{exe}
\ex Everyone loves someone.\pause
\ex Everyone loves someone, and that person is Billy.\pause
\ex Everyone loves someone. Don't pretend like you don't have someone special.\pause
\ex Someone is loved by everyone.\pause
\ex Someone is loved by everyone, and that person is Billy.\pause
\ex[??]{ Someone is loved by everyone. Don't pretend like you don't have someone special.}
\end{exe}
\end{frame}

\begin{frame}
	\frametitle{The Payoffs}\pause

	\begin{itemize}
		\item Both Players receive a payoff when the sentence is correctly communicated, represented by $x$.\pause
		\item If the more marked inverse scope is employed, both players suffer a slightly diminished payoff. We we refer to this amount as $i$.\pause
		\item If the Speaker employs passive voice, he suffers a slight loss $p$.\pause
		\item $|p+i|<|x|$ That is, even if we have to passivize and get inverse scope interpretation, it's always most preferable to get the intended interpretation.
	\end{itemize}

\end{frame}

\begin{frame}
\frametitle{The Decision Tree}
\begin{figure}
\begin{forest} 
for tree={grow=east,draw=black,line width=0.2pt,parent anchor=east,child anchor=west,edge={draw=black},edge label={\Huge\color{black}},edge path={\noexpand\path[\forestoption{edge}](!u.parent anchor) -- ([xshift=-1.6cm].child anchor) --    
      (.child anchor)\forestoption{edge label};
  },
  l sep=2cm,
} 
[Nature,rectangle, s sep=25pt,
  [Speaker,edge label={node[Below]{$\exists>\forall$}}
    [Hearer,edge label={node[Below]{Passive}}
	[{$-p-i,-i$},edge label={node[Below]{Inverse}}]
	[{$x-p,x$},edge label={node[Below]{Surface}}]
	]
    [Hearer,edge label={node[Above]{Active}}
	[{$x-i,x-i$},edge label={node[Below]{Inverse}}]
	[{$0,0$},edge label={node[Below]{Surface}}]
	]
  ]
  [Speaker,edge label={node[Above]{$\forall>\exists$}}
    [Hearer,edge label={node[Below]{Passive}}
	[{$x-p-i,x-i$},edge label={node[Below]{Inverse}}]
	[{$-p,0$},edge label={node[Below]{Surface}}]
	]
    [Hearer,edge label={node[Above]{Active}}
	[{$-i,-i$},edge label={node[Below]{Inverse}}]
	[{$x,x$},edge label={node[Below]{Surface}}]
	]
  ]
]
\end{forest}
\caption{Decision Flow of the Game of ``Everybody loves somebody''}
\end{figure}
\end{frame}

\begin{frame}[fragile]
\frametitle{Matrix for when Nature chooses $\forall > \exists$}
\begin{center}
\begin{figure}
\begin{tikzpicture}
\matrix[matrix of math nodes,every odd row/.style={align=right},every even row/.style={align=left},every node/.style={text width=1.5cm},row sep=0.2cm,column sep=0.2cm] (m) {
$x$&$-y$\\
$x$&$0$\\
$-i$&$x-p-i$\\
$-i$&$x-i$\\
};
\draw (m.north east) rectangle (m.south west);
\draw (m.north) -- (m.south);
\draw (m.east) -- (m.west);
\coordinate (a) at ($(m.north west)!0.25!(m.north east)$);
\coordinate (b) at ($(m.north west)!0.75!(m.north east)$);
\node[above=5pt of a,anchor=base] {Active};
\node[above=5pt of b,anchor=base] {Passive};
\coordinate (c) at ($(m.north west)!0.25!(m.south west)$);
\coordinate (d) at ($(m.north west)!0.75!(m.south west)$);
\node[left=2pt of c,text width=1cm]  {Surface};
\node[left=2pt of d,text width=1cm]  {Inverse};
\node[above=18pt of m.north] (firm b) {Speaker};
\node[left=1.6cm of m.west,rotate=90,align=center,anchor=center] {Hearer};
\end{tikzpicture}
\caption{Decision Flow of the Game of ``Everybody loves somebody''}
\end{figure}
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Matrix for when Nature chooses $\exists > \forall$}
\begin{center}
\begin{figure}
\begin{tikzpicture}
\matrix[matrix of math nodes,every odd row/.style={align=right},every even row/.style={align=left},every node/.style={text width=1.5cm},row sep=0.2cm,column sep=0.2cm] (m) {
$0$&$x-p$\\
$0$&$x$\\
$x-i$&$-p-i$\\
$x-i$&$-i$\\
};
\draw (m.north east) rectangle (m.south west);
\draw (m.north) -- (m.south);
\draw (m.east) -- (m.west);
\coordinate (a) at ($(m.north west)!0.25!(m.north east)$);
\coordinate (b) at ($(m.north west)!0.75!(m.north east)$);
\node[above=5pt of a,anchor=base] {Active};
\node[above=5pt of b,anchor=base] {Passive};
\coordinate (c) at ($(m.north west)!0.25!(m.south west)$);
\coordinate (d) at ($(m.north west)!0.75!(m.south west)$);
\node[left=2pt of c,text width=1cm]  {Surface};
\node[left=2pt of d,text width=1cm]  {Inverse};
\node[above=18pt of m.north] (firm b) {Speaker};
\node[left=1.6cm of m.west,rotate=90,align=center,anchor=center] {Hearer};
%\node[above=5pt of firm b]  {If Nature chooses $\forall>\exists$ in a scrambling language};
\end{tikzpicture}
\caption{Decision Flow of the Game of ``Everybody loves somebody''}
\end{figure}
\end{center}
\end{frame}
\begin{frame}
\frametitle{Scope in Scrambling Languages}\pause
\begin{itemize}
\item English has relatively rigid word order (subject-verb-object), but many languages have what is called `scrambling' which is free linear movement of nouns without the cost of transformations.\pause
\item Scope is systematically different in languages like these.\pause
\end{itemize}
\begin{exe}
\ex {\gll Har d\=aneshjui ye kit\=abi-ro mixune.\\
all student a book-OBJ reads\\
\trans{``Every student is reading a book.''}}\pause
\ex {\gll Ye ket\=abi-ro har d\=aneshjui mixune.\\
a book-OBJ all student reads\\
\trans{``Every student is reading a book.''}}
\end{exe}\pause
\begin{itemize}
\item However, both of these sentences \emph{must have} \textbf{surface scope}. They cannot be ambiguous.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{A Game Theoretic Account}\pause
\begin{itemize}
    \item Given our previous suggested constraints, we can predict these scope availabilities.\pause
    \item Remember, \textbf{surface scope} is preferred and \textbf{transformations} are costly.
\end{itemize}
\end{frame}

\begin{frame}
\begin{figure}

	\footnotesize
\begin{forest} 
for tree={grow=east,draw=black,line width=0.2pt,parent anchor=east,child anchor=west,edge={draw=black},edge label={\Huge\color{black}},edge path={\noexpand\path[\forestoption{edge}](!u.parent anchor) -- ([xshift=-1.6cm].child anchor) --    
      (.child anchor)\forestoption{edge label};
  },
  l sep=2cm,
} 
[Nature,rectangle, s sep=25pt,
  [Speaker,edge label={node[Below]{$\exists>\forall$}}
    [Hearer,edge label={node[Below]{Passive}}
	[{$-y-w,-w$},edge label={node[Below]{Inverse}}]
	[{$x-y,x$},edge label={node[Below]{Surface}}]
	]
    [Hearer,edge label={node[Above]{Active}}
	[{$x-w,x-w$},edge label={node[Below]{Inverse}}]
	[{$0,0$},edge label={node[Below]{Surface}}]
	]
[Hearer, edge label={node[Above]{Scramble}}
[{$-w,-w$}, edge label={node[Below]{Inverse}}]
[{$x,x$}, edge label={node[Below]{Surface}}]
]
  ]
  [Speaker,edge label={node[Above]{$\forall>\exists$}}
    [Hearer,edge label={node[Below]{Passive}}
	[{$x-y-w,x-w$},edge label={node[Below]{Inverse}}]
	[{$-y,0$},edge label={node[Below]{Surface}}]
	]
    [Hearer,edge label={node[Above]{Active}}
	[{$-w,-w$},edge label={node[Below]{Inverse}}]
	[{$x,x$},edge label={node[Below]{Surface}}]
	]
[Hearer, edge label={node[Above]{Scramble}}
[{$x-w$}, edge label={node[Below]{Inverse}}]
[{$0,0$}, edge label={node[Below]{Surface}}]
]
  ]
]
\end{forest}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{In an English-like language\ldots}\pause
\begin{itemize}
\item As assumed speakers \emph{want} to interpret quantifiers in linear order.\pause
\item When a speaker produces a costly transformation (like a passive) the listener assumes that the new surface word order is the intended scope order.\pause
\item If a speaker produces an untransformed sentence, the listener has two possible hypotheses: (1) the speaker intended surface scope, or (2) that the speaker intended inverse scope, but didn't want to undergo a costly transformation.\pause
\item These two possibilities produce scope ambiguity.
\end{itemize}
\end{frame}


\begin{frame}
\frametitle{In Scrambling Languages}\pause
\begin{itemize}
\item In scrambling languages, since speakers have greater flexibility in ordering, listeners make different assumptions about intended scope.\pause
\item If the speaker wants the object to scope over the subject, he can easily scramble it leftward.\pause
\item Since he can do this, the unscrambled sentence has an unambiguous surface scope interpretation.\pause
\item \textbf{Sidenote:} Potentially related, languages with scrambling/flexible word order, usually rely on things like passivization less often.
\end{itemize}
\end{frame}


\begin{frame}
\frametitle{Just a random difference?}\pause
\begin{itemize}
\item In addition to this correlation between rigid word-order and scrambling languages, we see that this theory still hold in rigid constructions in scrambling languages.\pause
\item In Persian, for example, although nouns are flexible, negation must always be on the same part of a verb.\pause
\item We should expect negative quantifiers to work similar to English sentences in that they produce ambiguity. This is the case:\pause
\begin{exe}
\ex {\gll Billy ye ket\=abi-ro na-xund.\\
Billy a book-OBJ not-read\\
\trans{``Billy didn't read a (particular) book.'' ($\exists>\neg$)or ``Billy didn't read any book.'' ($\neg>\exists$)}}
\end{exe}\pause
\item This holds in similar languages with scrambling and stable negation location (e.g. Korean).

\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Rigidity = Ambiguity; Flexiblity = Unambiguousness}

	\begin{itemize}
		\item The general theorem that arises from this analysis is that \emph{wherever} we have syntactic flexibility, we have ambiguity (and \textit{vice versa}.)
		\item This difference, in agreement with our theory, is true \emph{across constructions}, not necessarily \emph{across languages}.
		\item ``Scrambling'' languages are unambiguous in normal sentences, but are in more rigid constructions, ambiguity arises.
			\begin{itemize}
				\item This is because the ambiguity is not a language-specific parameter, but a result of the strategies employable i any given context.
			\end{itemize}
	\end{itemize}

\end{frame}

\begin{frame}
	\frametitle{Local Rigidity}\pause
	In scrambling languages, generally we have syntactic flexibility accompanied by unambiguous surface scope.\pause
	
\begin{exe}
\ex \begin{xlist}\label{chin}
\ex[]{\gll Meigeren dou zhuazou yige {n\"uren}.\\
everyone all arrest a woman\\
	\trans{``Everyone arrested a woman.'' (${\forall}>{\exists} $)\label{chin1}}\pause
}
\ex[]{\gll (You) yige {n\"uren} meigeren dou zhuazou.\\
(have) a woman everyone all arrest.\\
	\trans{``A woman was arrested by everyone.'' (${\exists}>{\forall}$)\label{chin2}}
}\end{xlist}
\end{exe}\pause

But in syncactically inflexible constructions, ambiguity arises.\pause

\begin{exe}
\ex{ \begin{xlist}
\ex[]{\gll Meigeren dou bei yige {n\"uren} zhuazou.\\
everyone all PASS a woman arrest\\
	\trans{``Everyone was arrested by a woman.'' ($\forall > \exists$, ${\exists} > {\forall}$)\label{chin3}}\pause
}
\ex[*]{Bei yige {n\"uren} meigeren dou zhuazou.\\
PASS a woman everyone all arrest\\
\label{chin4}
}\end{xlist}}
\end{exe}
\end{frame}


\begin{frame}
\frametitle{Goals and Intuitions}\pause
	\begin{itemize}
		\item Replace generative notions of syntactically-determined quantifier scope ambiguities with more plausible, externally-driven factors.
		\item Unify this account with other scope alternations (say, the unavailability of semantically implausible scope interpretations) into a general theory of scope where possible interpretations are \emph{pruned}, rather than derived by some syntactic engine.
	\end{itemize}
\end{frame}

\end{document}
